---
title: "get_plot_microbiome"
author: "Wanting Jin"
date: "9/2/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}
### get data for microbiome, Bile Acids, Tryptophan and Mood
dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'
setwd(dir1)

library(knitr)
library(ggplot2)
library(reshape2)
library(lme4)
source('data_analysis_MBML.R')
library(tidyr)  
library(stringr)
library(dplyr)
library(lmerTest)
library(magrittr)
library(qwraps2)
library(rstatix)
library(ggpubr)

#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results/',dir2)

filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
dat_trp = read.csv(filein_trp)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)

setwd(dir2)
filein_mood = './Data/processed_data/MPPD_Data_Master_09-13-2020.csv'
dat_mood = read.csv(filein_mood)

edataNames_BA = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')
edataNames_Trp = c('Tryptophan','Kynurenine','Kynurenic.Acid','Neopterin','Serotonin', 'KYN_TRP_Ratio', 'KA_KYN_Ratio')
edataNames_div = c(paste("wunifrac.PC.",1:4,sep='')
                  ,"pielou_e","observed_otus","faith_pd","shannon")

```

##Spaghetti plot of GAD7 and EPDS scores by visits grouped by psychitric history
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}
dat_GAD = dat_mood[1:30,c(1,21,48,50,52)]
dat_EPDS = dat_mood[1:30,c(1,21,54,57,60)]
colnames(dat_GAD)= c('Subject','Group','V1','V2','V3')
colnames(dat_EPDS)= c('Subject','Group','V1','V2','V3')


mx_GAD = melt(dat_GAD,measure.vars=c('V1','V2','V3'),id = c('Subject','Group'),variable.name = 'Visit',value.name = 'GAD7')
mx_EPDS = melt(dat_EPDS,measure.vars=c('V1','V2','V3'),id = c('Subject','Group'),variable.name = 'Visit',value.name = 'EPDS')
mx_mood = merge(mx_GAD,mx_EPDS,by = c('Subject','Group','Visit'))
mx_mood$nVisit = as.numeric(substring(mx_mood$Visit,2,2))
mx_mood$EPDS = as.numeric(mx_mood$EPDS)
mx_mood$GAD7 = as.numeric(mx_mood$GAD7)
mx_mood$Subject = substring(mx_mood$Subject,5,6)

his_labs = c("Anx(n = 7)","AnxorMDD(n = 9)","MDD(n = 4)","None(n = 10)")
names(his_labs) = c("Anx","AnxorMDD","MDD","None")

for(edataName in c('GAD7','EPDS')){
    p = ggplot(mx_mood)+
    facet_grid(.~Group)+
    geom_line(aes(nVisit,get(edataName),color = Subject))+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('scores')+
    ggtitle(edataName)+
    theme_classic()
    #ggsave(sprintf('%s/Report/plots/Figure2&3b trajectory/%s_goup_spaghetti.png',dir2,edataName),width = 6,height = 4)
    print(p)
}

for(edataName in c('GAD7','EPDS')){
    p = ggplot(mx_mood)+
    geom_line(aes(nVisit,get(edataName),fill = Subject,color = Group))+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('scores')+
    ggtitle(edataName)+
    theme_classic()
    #ggsave(sprintf('%s/Report/plots/Figure2&3b trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
    print(p)
}



```

## Boxplot of Tryptophan related pathway by visits
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=12, fig.height=6}


dat_trp1 = dat_trp[,c(1:2,44,46:51)]
mx_trp = melt(dat_trp1,measure.vars = edataNames_Trp,id = c("Subject","Visit"),variable.name = "Trp")

##calculate the repeated anova test among Visits

anovap = matrix(NA,7,1)
for(i in 1:length(edataNames_Trp)){
  anovap[i] <- anova_test(data = mx_trp[mx_trp$Trp==edataNames_Trp[i],],dv = value, wid = Subject,within = Visit)$ANOVA$p
}


#Boxplot of Tryptophan related pathway by visits with group comparison results
my_comparisons = list(c("V1","V2"),c("V2","V3"),c("V1","V3"))

rm.subject = c("16","22","28") #remove subjects lost to follow up
mx_trp1 = mx_trp[!mx_trp$Subject %in% rm.subject,]

ggboxplot(mx_trp1,x = "Visit", y = "value",fill = "Visit",outlier.shape = NA)+
  facet_wrap(~Trp, scales ="free_y",nrow = 1)+
  geom_point(aes(Visit,value,shape = Visit),position = position_jitter(width = 0.2),)+
  stat_boxplot(aes(Visit,value, fill = Visit),geom ='errorbar')+
  stat_compare_means(comparisons = my_comparisons,p.adjust.method = "bonferroni",paired = T)+
#  stat_compare_means(method = "anova",label = "p")+
  scale_fill_brewer(palette = "Set2")+
  xlab("Tryptophan related pathways")+
  ylab('ng/ml')+
  theme_classic()

```


## Boxplot of Bile Acids by visits
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=12, fig.height=12}



edataNames = edataNames_BA
#par(mfrow=c(3,3))
dat_ba1 = dat_ba[,3:21]
mx_ba = melt(dat_ba1,measure.vars = edataNames_BA,id = c("Subject","Visit"),variable.name = "BileAcids")
rm.subject = c("16","22","28")
mx_ba = mx_ba[!mx_ba$Subject %in% rm.subject,]


##boxplot of bile Acids by visits
ggplot(mx_ba,position = position_dodge(0.3))+
  geom_boxplot(aes(BileAcids,value,fill = Visit))+
  geom_point(aes(BileAcids,value,shape = Visit), position = position_jitterdodge(dodge.width = 0.8,jitter.width = 0.1))+
  stat_boxplot(aes(BileAcids,value, fill = Visit),geom ='errorbar')+
  scale_color_brewer(palette = "Set2")+
  scale_fill_brewer(palette = "Set2")+
  xlab("Bile Acids")+
  ylab('ng/ml')+
  theme_classic()


##boxplot of bile Acids by visits with group comparison results
my_comparisons = list(c("V1","V2"),c("V2","V3"),c("V1","V3"))
ggboxplot(mx_ba,x = "Visit", y = "value",fill = "Visit",outlier.shape = NA)+
  facet_wrap(~BileAcids,ncol = 9)+
  geom_point(aes(Visit,value,shape = Visit),position = position_jitter(width = 0.2),)+
  stat_boxplot(aes(Visit,value, fill = Visit),geom ='errorbar')+
  stat_compare_means(comparisons = my_comparisons,p.adjust.method = "bonferroni",paired = T)+
#  geom_text(data = plabel,aes(x = -Inf,-Inf,label = plabel),inherit.aes = FALSE)+
#  stat_compare_means(method = "anova",label = "p")+
    scale_fill_brewer(palette = "Set2")+
  xlab("Tryptophan related pathways")+
  ylab('ng/ml')+
  theme_classic()
  

```


##Boxplot of Tryptophan and Bile Acids by visits grouped by psychitric history
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}

his = matrix(NA,nrow = 30, ncol = 2)
colnames(his) = c("Subject","his")
his[,1]=c(1:30)
his[c(8,10,14,18,24:25,29),2] = 'Anx'
his[c(9,13,20,27),2] = 'MDD'
his[c(5:7,11:12,17,19,22,30),2] = 'AnxorMDD'
his[c(1:4,15:16,21,23,26,28),2] = 'None'

mx1 = merge(dat_trp1,his,by = 'Subject')
mx1 = na.omit(mx1)



for(edataName in edataNames_Trp){
    p = ggplot(mx1,position = position_dodge(width = 0.8))+
    stat_boxplot(aes(his,get(edataName), fill = Visit),geom ='errorbar')+
    geom_boxplot(aes(his,get(edataName),fill = Visit))+
    scale_x_discrete(name = "psychiatric history",labels = c("Anx"="Anx(n = 7)","MDD"="MDD(n = 4)","AnxorMDD"="AnxorMDD(n = 9)","None"="None(n = 10)"))+
    scale_fill_brewer(palette = "Set2")+
    scale_color_brewer(palette = "Set2")+
    xlab("Psychiatric History")+
    ylab('ng/ml')+
    ggtitle(edataName)+
    theme_classic()+
    #ggsave(sprintf('%s/Report/plots/trajectory/%s_boxplot.png',dir2,edataName),width = 6,height = 4)
    print(p)
}

mx1 = merge(dat_ba1,his,by = 'Subject')
for(edataName in edataNames_BA){
    p = ggplot(mx1,position = position_dodge(width = 0.8))+
    stat_boxplot(aes(his,get(edataName), fill = Visit),geom ='errorbar')+
    geom_boxplot(aes(his,get(edataName),fill = Visit))+
    scale_x_discrete(name = "psychiatric history",labels = c("Anx"="Anx(n = 7)","MDD"="MDD(n = 4)","AnxorMDD"="AnxorMDD(n = 9)","None"="None(n = 10)"))+
    scale_fill_brewer(palette = "Set2")+
    scale_color_brewer(palette = "Set2")+
    xlab("Psychiatric History")+
      
    ylab('ng/ml')+
    ggtitle(edataName)+
    theme_classic()+
    #ggsave(sprintf('%s/Report/plots/trajectory/%s_boxplot.png',dir2,edataName),width = 6,height = 4)
    print(p)
}



```


##Spaghetti plot of Tryptophan and Bile Acids by visits grouped by psychitric history
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=12, fig.height=18}

mx1 = merge(dat_trp1,his,by = 'Subject')
mx1 = na.omit(mx1)
mx1$nVisit = as.numeric(substring(mx1$Visit,2,2))
mx1$Subject = as.factor(mx1$Subject)
#edataName = edataNames[1]

his_labs = c("Anx(n = 7)","AnxorMDD(n = 9)","MDD(n = 4)","None(n = 10)")
names(his_labs) = c("Anx","AnxorMDD","MDD","None")

#spaghetti plot of Tryptophan by visit
par(mfrow=c(2,2))
edataname_trp = c('Tryptophan','Neopterin')
for(edataName in edataname_trp){

    p = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,get(edataName),color = Subject))+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle(edataName)+
    theme_classic()
#    ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
    print(p)
}

#spaghetti plot of Bile Acids by visit

mx1 = merge(dat_ba1,his,by = 'Subject')
mx1$nVisit = as.numeric(substring(mx1$Visit,2,2))
mx1$Subject = as.factor(mx1$Subject)
for(edataName in edataNames_BA){
    p = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,get(edataName),color = Subject))+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle(edataName)+
    theme_classic()+
#    ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
    print(p)
}


##Grouping the spagetti plots of tryptophan and neopterin
    pa = ggplot(mx_mood)+
    facet_grid(.~Group,labeller = labeller(Group = his_labs))+
    geom_line(aes(nVisit,GAD7,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('GAD7')+
    theme(legend.position="none")+
    theme_classic()
    pb = ggplot(mx_mood)+
    facet_grid(.~Group,labeller = labeller(Group = his_labs))+
    geom_line(aes(nVisit,EPDS,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('EPDS')+
    theme(legend.position="none")+
    theme_classic()


    p1 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Tryptophan,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Tryptophan')+
    theme(legend.position="none")+
    theme_classic()
    p2 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Kynurenine,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Kynurenine')+
    theme(legend.position="none")+
    theme_classic()
    p3 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Kynurenic.Acid,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Kynurenic.Acid')+
    theme(legend.position="none")+
    theme_classic()
    p4 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Neopterin,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Neopterin')+
    theme(legend.position="none")+
    theme_classic()
    p5 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,KA_KYN_Ratio,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('KA_KYN_Ratio')+
    theme(legend.position="none")+
    theme_classic()
    p6 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,KYN_TRP_Ratio,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('KYN_TRP_Ratio')+
    theme(legend.position="none")+
    theme_classic()


    library(gridExtra)
    grid.arrange(pa,pb,p1,p2,p3,p4,p5,p6,ncol = 2)


##Grouping the boxplots of tryptophan and neopterin
    pa=ggplot(mx_mood,position = position_dodge(width = 0.8))+
    stat_boxplot(aes(his,GAD7, fill = Visit),geom ='errorbar')+
    geom_boxplot(aes(his,GAD7,fill = Visit))+
    scale_x_discrete(name = "psychiatric history",labels = c("Anx"="Anx(n = 7)","MDD"="MDD(n = 4)","AnxorMDD"="AnxorMDD(n = 9)","None"="None(n = 10)"))+
    scale_fill_brewer(palette = "Set2")+
    scale_color_brewer(palette = "Set2")+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('GAD7')+
    theme(legend.position="none")+
    theme_classic()
    pb=ggplot(mx_mood,position = position_dodge(width = 0.8))+
    stat_boxplot(aes(his,EPDS, fill = Visit),geom ='errorbar')+
    geom_boxplot(aes(his,EPDS,fill = Visit))+
    scale_x_discrete(name = "psychiatric history",labels = c("Anx"="Anx(n = 7)","MDD"="MDD(n = 4)","AnxorMDD"="AnxorMDD(n = 9)","None"="None(n = 10)"))+
    scale_fill_brewer(palette = "Set2")+
    scale_color_brewer(palette = "Set2")+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('EPDS')+
    theme(legend.position="none")+
    theme_classic() 

    p1 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Tryptophan,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Tryptophan')+
    theme(legend.position="none")+
    theme_classic()
    p2 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Kynurenine,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Kynurenine')+
    theme(legend.position="none")+
    theme_classic()
    p3 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Kynurenic.Acid,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Kynurenic.Acid')+
    theme(legend.position="none")+
    theme_classic()
    p4 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,Neopterin,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('Neopterin')+
    theme(legend.position="none")+
    theme_classic()
    p5 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,KA_KYN_Ratio,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('KA_KYN_Ratio')+
    theme(legend.position="none")+
    theme_classic()
    p6 = ggplot(mx1)+
    facet_grid(.~his,labeller = labeller(his = his_labs))+
    geom_line(aes(nVisit,KYN_TRP_Ratio,color = Subject),show.legend = FALSE)+
    scale_x_continuous(breaks = seq(1,3,1))+
    xlab("Visit")+
    ylab('ng/ml')+
    ggtitle('KYN_TRP_Ratio')+
    theme(legend.position="none")+
    theme_classic()


    library(gridExtra)
    grid.arrange(pa,pb,p1,p2,p3,p4,p5,p6,ncol = 2)    
  
    
    




```


##plot of change of Tryptophans vs Bile Acids at V1
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}
## reshape the responses data to wide format

mx2= merge(dat_trp1,dat_ba1,by = c('Subject','Visit'))
mx2_long = reshape(mx2, idvar='Subject', timevar = 'Visit', direction='wide')

edataNames_diff21 = c()
edataNames_diff32 = c()

# calculate the diff of Trp pathway variables
for(edataName in edataNames_Trp){
  edataName_visits = paste(edataName,c('V1','V2','V3'),sep='.')
  edataName_diff21 = paste(edataName,'V21',sep='.')
  edataName_diff32 = paste(edataName,'V32',sep='.')
  mx2_long[[edataName_diff21]] = mx2_long[, edataName_visits[2]] - mx2_long[, edataName_visits[1]]
  mx2_long[[edataName_diff32]] = mx2_long[, edataName_visits[3]] - mx2_long[, edataName_visits[2]]
  
  edataNames_diff21 = c(edataNames_diff21, edataName_diff21)
  edataNames_diff32 = c(edataNames_diff32, edataName_diff32)

}

edataName = c('Tryptophan.V21','Kynurenine.V21',rep('Neopterin.V21',7),rep('KA_KYN_Ratio.V21',3))
cdataName = c(rep('TUDCA.V1',2),'TaMCA.V1','THCA.V1','TUDCA.V1','TCA.V1','GCA.V1','TCDCA.V1','TDCA.V1','TUDCA.V1','TCA.V1','GCA.V1')

par(mfrow=c(4,3))
for(i in 1:12){
  plot(mx2_long[,cdataName[i]],mx2_long[,edataName[i]],xlab=cdataName[i],ylab=edataName[i],cex.lab = 1.5)
  if(class(mx2_long[,cdataName[i]]) == 'numeric'){
    if(var(mx2_long[,cdataName[i]], na.rm = TRUE) !=0){
      #abline(lm1)
      abline(lm(mx2_long[,edataName[i]] ~ mx2_long[,cdataName[i]]))
    }
  }
}
```



##plot of Bile Acids vs Diversity
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}
## reshape the responses data to wide format

mx2= merge(dat_trp1,dat_ba1,by = c('Subject','Visit'))


edataName = c('GUDCA','UDCA')
cdataName = c("pielou_e","observed_otus","faith_pd","shannon")
  
```

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  setwd(dir1)

  filein_div = sprintf('%s/qiime2_diversity.csv',dir_mb)
  dat_div = read.csv(filein_div)
  mx3 = merge(dat_trp[,c(1:5,44:51)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)
  mx3 = merge(mx3,dat_div,by='barcode')


  edataNames = c('GUDCA','UDCA')
  cdataNames = c("pielou_e","observed_otus","faith_pd","shannon")
  clabNames = c("Pielou's evenness","Observed OTUs","Faith's PD","Shannon")
  timeName = 'Visit'
  
plot_longitudinal = function(mx2,cdataNames,edataNames, timeName
                                         , dataName='out',dir_out = 'output',y0,p) {
  timeNames = unique(mx2[,timeName])
  n_time = length(timeNames)
  #cols = ('red','green','blue','yellow','pink','black')
  cols = rainbow(n_time)
  for(j in 1:length(edataNames)){
    for(i in 1:length(cdataNames)){
        x_min = min(0,mx2[,cdataNames[i]], na.rm= TRUE)
        x_max = max(mx2[,cdataNames[i]], na.rm = TRUE)
        y = mx2[,edataNames[j]]
        x = mx2[,cdataNames[i]]
        idx_keep = which(!is.na(x) & !is.na(y))
        idx_timeName = match(mx2[,timeName], timeNames)
       plot(x[idx_keep], y[idx_keep]
          , pch = idx_timeName[idx_keep]
          , col = cols[idx_timeName[idx_keep]]
          , xlab=clabNames[i], ylab=edataNames[j],cex.lab = 1.5
          )
        legend('topright', legend=timeNames, col=cols, pch=1:n_time)
        abline(lm(y~x))
        xmin = min(mx2[,cdataNames[i]], na.rm= TRUE)
        xmax = max(mx2[,cdataNames[i]], na.rm= TRUE)
        text(xmin+0.15*(xmax-xmin),y0,label = paste('p=',p[i]),cex = 1.5)
    }
  }
}
p1 = c(0.0016,0.0268,0.0025,0.0019)
p2 = c("5.77E-07","4.70E-05","6.51E-06","1.15E-07")
p3 = 0.0011
  par(mfrow=c(4,3))
  plot_longitudinal(mx3,cdataNames,'GUDCA', timeName,dataName = dataName, dir_out = dir_out,y0 = 2.1,p = p1)
  plot_longitudinal(mx3,cdataNames,'UDCA', timeName,dataName = dataName, dir_out = dir_out,y0 = -0.9,p = p2)
  plot_longitudinal(mx3,'pielou_e', 'CDCA',timeName,dataName = dataName, dir_out = dir_out,y0 = 1.1, p = p3)


```

