---
title: "get_data_microbiome_qiime2"
author: "Wanting Jin"
date: "6/3/2020"

output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Outline of the analysis
* PCA of alpha-diversity data from all three visits
* linear mixed regression analysis between alpha-diversity and metabolites
* Transform data by log to normalize the data
* Boxplot and histogram to analyze the distribution and outliers
* Summary statistics to summarize the data
* Correlation heatmap between variables




```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
### load data file Microbiome diversity raw file
dir1 = 'E:/Github/mbml'
dir2 = 'E:/Wanting/mbml'
setwd(dir2)

library(knitr)
library(ggplot2)
library(reshape2)
library(lme4)
library(tidyr)
library(stringr)
library(dplyr)
library(lmerTest)
library(magrittr)
library(qwraps2)

#dir_beha = 'behaviordata'
dir_mb = sprintf('%s/processed_data',dir1)
dir_raw = './Data/otus/qiime2/QZA/exported_data'
dir_out = sprintf('%s/results',dir2)


beta=read.table(sprintf("%s/unweighted_unifrac_pcoa_mod_results.txt",dir_raw),header=FALSE)
weight_beta=read.table(sprintf("%s/weighted_unifrac_pcoa_mod_results.txt",dir_raw),header=FALSE)

weight_beta=weight_beta[,c(1:5)]
colnames(weight_beta)=c("barcode","wunifrac.PC.1","wunifrac.PC.2","wunifrac.PC.3","wunifrac.PC.4")

beta=beta[,c(1:5)]
colnames(beta)=c("barcode","unifrac.PC.1","unifrac.PC.2","unifrac.PC.3","unifrac.PC.4")

  #div_beta = merge(weight_beta,beta,by=1)
  
div_beta = weight_beta

  #div_beta[,1]<-substr(div_beta[,1],0,5)
  
  
  
  ### alpha diversity
  mx_alpha = c()
  divNames = c('evenness','observed_otus','faith_pd','shannon')
  
  
  filein_alpha1= sprintf("%s/evenness.tsv",dir_raw)
  filein_alpha2= sprintf("%s/observed_otus.tsv",dir_raw)
  filein_alpha3= sprintf("%s/faith_pd.tsv",dir_raw)
  filein_alpha4= sprintf("%s/shannon.tsv",dir_raw)

  evenness=read.table(filein_alpha1, header=TRUE, na.strings = "NA",fill=TRUE)
  observed_otus=read.table(filein_alpha2, header=TRUE, na.strings = "NA",fill=TRUE)
  faith_pd=read.table(filein_alpha3, header=TRUE, na.strings = "NA",fill=TRUE)
  shannon=read.table(filein_alpha4, header=TRUE, na.strings = "NA",fill=TRUE)
  
  
  mx2 = cbind(evenness, observed_otus, faith_pd, shannon)
  barcode = rownames(mx2)
  mx3 = data.frame(barcode, mx2)
  
  
  ## merge alpha and beta diversity
  mx_div = merge(div_beta, mx3, by = 1)
  
  
  

#  fileout_div_cvrt = sprintf('%s/qiime2_diversity.csv',dir_mb)
#  write.csv(mx_div,fileout_div_cvrt,row.names = FALSE)


```





## PCA of alpha-diversity data from all three visits(plot by factoextra package)
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE}
dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'
setwd(dir1)
library(knitr)
library(ggplot2)
library(reshape2)
library(factoextra)
#edataNames = c("chao1","observed_species","PD_whole_tree","shannon")

pca_div <- princomp(mx3[,2:5],cor = T)#PCA of 4 alpha-diversity data by princomp function

fviz_eig(pca_div,addlabels = T)#Screenplot by using factoextra package

#Plot of accumulated percentage of explained variances
pca_sum <- summary(pca_div,loadings = T)
pca_sum$pva= pca_sum$sdev^2/sum(pca_sum$sdev^2)
for(i in 1:4){
  pca_sum$cpv[i] <- sum(pca_sum$pva[1:i])
  pca_sum$comp[i] = i;
}
pca_var <- data.frame(PC = pca_sum$comp, pvar = pca_sum$pva, cpvar = pca_sum$cpv)
screep <-ggplot(data = pca_var,aes(x = PC,y = cpvar))+
  geom_point()+
  geom_line()+
  geom_line(y = 0.95,linetype = 2)+
  geom_text(y = 0.96, x = 1,label = "0.95")+
  ylim(0.5,1)+
  labs(x= "PC", y = "Cumulative Percent of Variance")
screep



# Biplot
#fviz_pca_biplot(pca_div,col.ind = dat_div$Visit, addEllipses = T, label = "var", col.var = "black", reoek = TRUE, legend.title = "Visit")

#dat_div$Subject <- as.factor(dat_div$Subject)
#fviz_pca_biplot(pca_div,col.ind = dat_div$Subject, addEllipses = F, label = "var", col.var = "black", reoek = TRUE, legend.title = "Subject")

# Output the PCs of Alpha diversity

adiv <- predict(pca_div)
colnames(adiv) <- c("alphadiv.PC.1","alphadiv.PC.2","alphadiv.PC.3","alphadiv.PC.4")
mx_div <- cbind(mx_div,adiv)
fileout_div_cvrt = sprintf('%s/qiime2_diversity.csv',dir_mb)
write.csv(mx_div,fileout_div_cvrt,row.names = FALSE)

#fileout_data = sprintf('%s/mbml.csv',dir_mb)
#write.csv(dat_div,fileout_data, row.names = FALSE)

```

\newpage
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.height = 8, fig.width = 6}
#Plot of Loadings
divNames = c('evenness','observed_otus','faith_pd','shannon')

loadings <-cbind(divNames,melt(pca_div$loadings[,1:4]))
barplot1 <- ggplot(data = loadings)+
  geom_bar(aes(x = Var1, y = value),stat = "identity",width = 0.8)+
  facet_grid(Var2~.)+
  labs(title = "Loadings for PC1~PC4 of alpha-diversity", x = "alpha-diversity", y = "Loadings")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))
print(barplot1)

```



### Bile Acids of three visits as function of beta and alpha diversity of microbiome

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}

dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'


setwd(dir1)


library(knitr)
library("lme4")
library(lmerTest)
#library(survival)
#library(coxme)
source('data_analysis_MBML.R')





#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results/qiime2',dir2)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)
filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
dat_trp = read.csv(filein_trp)
filein_div = sprintf('%s/qiime2_diversity.csv',dir_mb)
dat_div = read.csv(filein_div)
mx2 = merge(dat_trp[,c(1:5,44:51)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)
mx2 = merge(mx2,dat_div,by='barcode')



edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')




dataset = 'BA'






cvrtNames = 'Visit'
randomName = 'Subject'


  
  cdataNames = c(paste("wunifrac.PC.",1:4,sep='')
                  ,"pielou_e","observed_otus","faith_pd","shannon")
  

  dataName = sprintf('%s_vs_div_q2',dataset)

  
  # Linear-mixed effect longitudinal model of random intercept and slope
  pairwise_association_test_lmm(mx2, cdataNames, edataNames, randomName, cvrtNames,dataName = dataName, dir_out = dir_out)
  


```




```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  
  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```




### Change of Bile Acids of three visits as function of V1/V2 of beta and alpha diversity of microbiome

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}

setwd(dir1)
library(lmerTest)

source('data_analysis_MBML.R')


#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results/qiime2',dir2)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)
filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
dat_trp = read.csv(filein_trp)
filein_div = sprintf('%s/qiime2_diversity.csv',dir_mb)
dat_div = read.csv(filein_div)
mx2 = merge(dat_trp[,c(1:5,44:51)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)
mx2 = merge(mx2,dat_div,by='barcode')


edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')

dataset = 'BA'

cvrtNames = 'Visit'
randomName = 'Subject'

cdataNames = c(paste("wunifrac.PC.",1:4,sep='')
                  ,"pielou_e","observed_otus","faith_pd","shannon")

  library(reshape2)
  #test1 = dcast(melt(mx2,id.vars=c('Subject','Visit')), Subject~variable+Visit)
  
  ## reshape the responses data to wide format
  mx2_long = reshape(mx2, idvar='Subject', timevar = 'Visit', direction='wide')
  
  edataNames_diff21 = c()
  edataNames_diff32 = c()
  
  # calculate the diff of Trp pathway variables
  for(edataName in edataNames){
    edataName_visits = paste(edataName,c('V1','V2','V3'),sep='.')
    edataName_diff21 = paste(edataName,'V21',sep='.')
    edataName_diff32 = paste(edataName,'V32',sep='.')
    mx2_long[[edataName_diff21]] = mx2_long[, edataName_visits[2]] - mx2_long[, edataName_visits[1]]
    mx2_long[[edataName_diff32]] = mx2_long[, edataName_visits[3]] - mx2_long[, edataName_visits[2]]
    
    edataNames_diff21 = c(edataNames_diff21, edataName_diff21)
    edataNames_diff32 = c(edataNames_diff32, edataName_diff32)

  }
  
  # association between div v1 and change of Trp
  dataName = sprintf('%s_diff_vs_div_q2_v1',dataset)
  cdataNames_v1 = paste(cdataNames, 'V1',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)
  
  
    # association between div v2 and change of Trp
  dataName = sprintf('%s_diff_vs_div_q2_v2',dataset)
  cdataNames_v1 = paste(cdataNames, 'V2',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)

```

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  
  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```




### Tryptophan of three visits as function of beta and alpha diversity of microbiome

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}

dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'

setwd(dir1)

library(knitr)
library("lme4")
library(lmerTest)
#library(survival)
#library(coxme)

source('data_analysis_MBML.R')


#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results/qiime2',dir2)
#filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
#mx2 = read.csv(filein_trp)


edataNames = c('Tryptophan','Kynurenine','Kynurenic.Acid','Neopterin','Serotonin', 'KYN_TRP_Ratio', 'KA_KYN_Ratio')
dataset = 'Trp'
cvrtNames = 'Visit'
randomName = 'Subject'
cdataNames = c(paste("wunifrac.PC.",1:4,sep='')
                ,"pielou_e","observed_otus","faith_pd","shannon")
dataName = sprintf('%s_vs_div_q2',dataset)

pairwise_association_test_lmm(mx2, cdataNames, edataNames, randomName, cvrtNames,dataName = dataName, dir_out = dir_out)

```




```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  
  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```




### Change of Tryptophan of three visits as function of V1/V2 of beta and alpha diversity of microbiome

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}

setwd(dir1)


library(knitr)
library("lme4")
library(lmerTest)
#library(survival)
#library(coxme)



source('data_analysis_MBML.R')


#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results/qiime2',dir2)


#filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
#mx2 = read.csv(filein_trp)

edataNames = c('Tryptophan','Kynurenine','Kynurenic.Acid','Neopterin','Serotonin', 'KYN_TRP_Ratio', 'KA_KYN_Ratio')
dataset = 'Trp'
cvrtNames = 'Visit'
randomName = 'Subject'
cdataNames = c(paste("wunifrac.PC.",1:4,sep='')
                  ,"pielou_e","observed_otus","faith_pd","shannon")
  

  library(reshape2)
  #test1 = dcast(melt(mx2,id.vars=c('Subject','Visit')), Subject~variable+Visit)
  
  ## reshape the responses data to wide format
  mx2_long = reshape(mx2, idvar='Subject', timevar = 'Visit', direction='wide')
  
  edataNames_diff21 = c()
  edataNames_diff32 = c()
  
  # calculate the diff of Trp pathway variables
  for(edataName in edataNames){
    edataName_visits = paste(edataName,c('V1','V2','V3'),sep='.')
    edataName_diff21 = paste(edataName,'V21',sep='.')
    edataName_diff32 = paste(edataName,'V32',sep='.')
    mx2_long[[edataName_diff21]] = mx2_long[, edataName_visits[2]] - mx2_long[, edataName_visits[1]]
    mx2_long[[edataName_diff32]] = mx2_long[, edataName_visits[3]] - mx2_long[, edataName_visits[2]]
    
    edataNames_diff21 = c(edataNames_diff21, edataName_diff21)
    edataNames_diff32 = c(edataNames_diff32, edataName_diff32)

  }
  
  # association between div v1 and change of Trp
  dataName = sprintf('%s_diff_vs_div_q2_v1',dataset)
  cdataNames_v1 = paste(cdataNames, 'V1',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)
  
  
    # association between div v2 and change of Trp
  dataName = sprintf('%s_diff_vs_div_q2_v2',dataset)
  cdataNames_v1 = paste(cdataNames, 'V2',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)

```

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  
  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```




