---
title: "get_analysis_ba_PCA"
author: "Wanting Jin"
date: "11/18/2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Outline of the analysis
* Principle Components analysis of processed Bile Acids data by visit1~3
* K-means clustering of Tryptophan and processed Bile Acids

## PCA of processed Bile Acids data by visit
\newpage
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.height = 8, fig.width = 6}
dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'
setwd(dir1)

library(knitr)
library(ggplot2)
library(reshape2)
library(lme4)
source('data_analysis_MBML.R')

dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_out = sprintf('%s/results',dir2)
filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)
filein_trp = sprintf('%s/mbml_trp0.csv',dir_mb)
dat_trp = read.csv(filein_trp)

filein_div = sprintf('%s/qiime2_diversity.csv',dir_mb)
dat_div = read.csv(filein_div)

edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')
setwd(dir2)
filein_mood = './Data/processed_data/MPPD_Data_Master_09-13-2020.csv'
dat_mood = read.csv(filein_mood)
dat_GAD = dat_mood[1:30,c(1,21,48,50,52)]
dat_EPDS = dat_mood[1:30,c(1,21,54,57,60)]
colnames(dat_GAD)= c('Subject','Group','V1','V2','V3')
colnames(dat_EPDS)= c('Subject','Group','V1','V2','V3')
mx_GAD = melt(dat_GAD,measure.vars=c('V1','V2','V3'),id = c('Subject','Group'),variable.name = 'Visit',value.name = 'GAD7')
mx_EPDS = melt(dat_EPDS,measure.vars=c('V1','V2','V3'),id = c('Subject','Group'),variable.name = 'Visit',value.name = 'EPDS')
mx_mood = merge(mx_GAD,mx_EPDS,by = c('Subject','Group','Visit'))
mx_mood$nVisit = as.numeric(substring(mx_mood$Visit,2,2))
mx_mood$EPDS = as.numeric(mx_mood$EPDS)
mx_mood$GAD7 = as.numeric(mx_mood$GAD7)
mx_mood$Subject = as.numeric(substring(mx_mood$Subject,5,6))


```



## K-means clustering of Bile Acids data from all three visits
\newpage
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE, figures-side, fig.show="hold", out.width="50%"}
dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'
setwd(dir1)
library(factoextra)



edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')
cluster_ba = dat_ba
cluster_ba$nVisit = as.numeric(substring(cluster_ba$Visit,2,2))


ks = 2:3
set.seed(20)



##converting to long format
mx1 = merge(dat_trp[,c(1:5,44:51)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)
mx2 = merge(mx1,dat_div,by='barcode')
mx3 = merge(mx2,mx_mood, by = c('Subject','Visit'))
ba_long = reshape(dat_ba, idvar='Subject', timevar = 'Visit', direction='wide')
ba_long = ba_long[-c(16,22,28),]
edataNames = c('UDCA','CDCA','TaMCA')


for(edataName in c('UDCA','CDCA','TaMCA')){
  edataName_visits = paste(edataName,c('V1','V2','V3'),sep='.')
  
  for(k in ks){
      clusters <- kmeans(ba_long[,edataName_visits],k)
      cluster_ba = ba_long
      cluster_ba$cluster = clusters$cluster
      for(i in 1:k){
        sprintf('k = ', i)
        print(cluster_ba[cluster_ba$cluster == i,1])
      }
      
      dat_cluster = merge(mx3,cluster_ba[,c(1,59)],by = 'Subject')
      dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
      
      #spaghetti plot of bile acids by clusters
      p = ggplot(dat_cluster)+
        facet_grid(.~cluster)+
        geom_line(aes(nVisit,get(edataName),color = as.factor(Subject)))+
        scale_x_continuous(breaks = seq(1,3,1))+
        #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
        xlab("Visit")+
        ylab('ng/ml')+
        ggtitle(edataName)+
        theme_classic()
        #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
      print(p)
      
      #spaghetti plot of alpha-diversity by clusters
      for(divname in c('shannon','pielou_e','GAD7','EPDS')){
        p = ggplot(dat_cluster)+
          facet_grid(.~cluster)+
          geom_line(aes(nVisit,get(divname),color = as.factor(Subject)))+
          scale_x_continuous(breaks = seq(1,3,1))+
          #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
          xlab("Visit")+
          ylab('ng/ml')+
          ggtitle(sprintf('%s clustered by %s ',divname, edataName))+
          theme_classic()
          #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
          print(p)
      }
      #dataMatrix = as.matrix(ba_long[,edataName_visits])
      #image(t(dataMatrix)[,order(cluster_ba$cluster)],yaxt = "n",main = "clustered Data")
  }
}


trp_long = reshape(dat_trp[,c(1:5,44:51)], idvar='Subject', timevar = 'Visit', direction='wide')
trp_long = trp_long[!trp_long$Subject %in% c(6,7,16,22,28),]
edataNames = c('Kynurenine','Kynurenic.Acid','Neopterin')


for(edataName in edataNames){
  edataName_visits = paste(edataName,c('V1','V2','V3'),sep='.')
  
  for(k in ks){
      clusters <- kmeans(trp_long[,edataName_visits],k)
      cluster_trp = trp_long
      cluster_trp$cluster = clusters$cluster
       for(i in 1:k){
        print(cluster_trp[cluster_trp$cluster == i,1])
      }
      dat_cluster = merge(mx3,cluster_trp[,c(1,35)],by = 'Subject')
      dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
      
      #spaghetti plot of bile acids by clusters
      p = ggplot(dat_cluster)+
        facet_grid(.~cluster)+
        geom_line(aes(nVisit,get(edataName),color = as.factor(Subject)))+
        scale_x_continuous(breaks = seq(1,3,1))+
        #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
        xlab("Visit")+
        ylab('ng/ml')+
        ggtitle(edataName)+
        theme_classic()
        #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
      print(p)
      
      #spaghetti plot of alpha-diversity by clusters
      for(divname in c('shannon','pielou_e','GAD7','EPDS')){
        p = ggplot(dat_cluster)+
          facet_grid(.~cluster)+
          geom_line(aes(nVisit,get(divname),color = as.factor(Subject)))+
          scale_x_continuous(breaks = seq(1,3,1))+
          #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
          xlab("Visit")+
          ylab('ng/ml')+
          ggtitle(sprintf('%s clustered by %s ',divname, edataName))+
          theme_classic()
          #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
          print(p)
      }
      #dataMatrix = as.matrix(ba_long[,edataName_visits])
      #image(t(dataMatrix)[,order(cluster_ba$cluster)],yaxt = "n",main = "clustered Data")
      
  }
}


```


## Combine K-means clustering of Bile Acids and tryptophan data from all three visits
\newpage
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE, figures-side, fig.width=12, fig.height=12}
dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'
setwd(dir1)
library(factoextra)



cluster_ba = dat_ba
cluster_ba$nVisit = as.numeric(substring(cluster_ba$Visit,2,2))


k = 2
set.seed(20)



##converting to long format
mx1 = merge(dat_trp[,c(1:5,40:47)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)
mx2 = merge(mx1,dat_div,by='barcode')
mx3 = merge(mx2,mx_mood, by = c('Subject','Visit'))
ba_long = reshape(dat_ba, idvar='Subject', timevar = 'Visit', direction='wide')
ba_long = ba_long[-c(16,22,28),]
edataNames = c('UDCA','CDCA','TaMCA')

cluster_ba = ba_long
edataName_visits = paste('UDCA',c('V1','V2','V3'),sep='.')
clusters <- kmeans(ba_long[,edataName_visits],k)
cluster_ba$cluster = clusters$cluster
dat_cluster = merge(mx3,cluster_ba[,c(1,59)],by = 'Subject')
dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
#spaghetti plot of bile acids by clusters
p1 = ggplot(dat_cluster)+
  facet_grid(.~cluster)+
  geom_line(aes(nVisit,UDCA,color = as.factor(Subject)))+
  scale_x_continuous(breaks = seq(1,3,1))+
  #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
  xlab("Visit")+
  ylab('ng/ml')+
  ggtitle('UDCA')+
  theme_classic()+
  labs(color = 'Subject')
  #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
  #print(p)

edataName_visits = paste('CDCA',c('V1','V2','V3'),sep='.')
clusters <- kmeans(ba_long[,edataName_visits],k)
cluster_ba$cluster = clusters$cluster
dat_cluster = merge(mx3,cluster_ba[,c(1,59)],by = 'Subject')
dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
p2 = ggplot(dat_cluster)+
  facet_grid(.~cluster)+
  geom_line(aes(nVisit,CDCA,color = as.factor(Subject)))+
  scale_x_continuous(breaks = seq(1,3,1))+
  #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
  xlab("Visit")+
  ylab('ng/ml')+
  ggtitle('CDCA')+
  theme_classic()+
  labs(color = 'Subject')
  #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
  #print(p)

edataName_visits = paste('TaMCA',c('V1','V2','V3'),sep='.')
clusters <- kmeans(ba_long[,edataName_visits],k)
cluster_ba$cluster = clusters$cluster
dat_cluster = merge(mx3,cluster_ba[,c(1,59)],by = 'Subject')
dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
p3 = ggplot(dat_cluster)+
  facet_grid(.~cluster)+
  geom_line(aes(nVisit,TaMCA,color = as.factor(Subject)))+
  scale_x_continuous(breaks = seq(1,3,1))+
  #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
  xlab("Visit")+
  ylab('ng/ml')+
  ggtitle('TaMCA')+
  theme_classic()+
  labs(color = 'Subject')
  #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
  #print(p)


trp_long = reshape(dat_trp[,c(1:5,40:47)], idvar='Subject', timevar = 'Visit', direction='wide')
trp_long = trp_long[!trp_long$Subject %in% c(16,22,28),]
edataNames = c('Kynurenine','Kynurenic.Acid','Neopterin')
cluster_trp = trp_long


edataName_visits = paste('Kynurenine',c('V1','V2','V3'),sep='.')
clusters <- kmeans(trp_long[,edataName_visits],k)
cluster_trp$cluster = clusters$cluster
dat_cluster = merge(mx3,cluster_trp[,c(1,35)],by = 'Subject')
dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
#spaghetti plot of bile acids by clusters
p4 = ggplot(dat_cluster)+
  facet_grid(.~cluster)+
  geom_line(aes(nVisit,Kynurenine,color = as.factor(Subject)))+
  scale_x_continuous(breaks = seq(1,3,1))+
  #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
  xlab("Visit")+
  ylab('ng/ml')+
  ggtitle('Kynurenine')+
  theme_classic()+
  labs(color = 'Subject')
  #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
  #print(p)

edataName_visits = paste('Kynurenic.Acid',c('V1','V2','V3'),sep='.')
clusters <- kmeans(trp_long[,edataName_visits],k)
cluster_trp$cluster = clusters$cluster
dat_cluster = merge(mx3,cluster_trp[,c(1,35)],by = 'Subject')
dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
#spaghetti plot of bile acids by clusters
p5 = ggplot(dat_cluster)+
  facet_grid(.~cluster)+
  geom_line(aes(nVisit,Kynurenic.Acid,color = as.factor(Subject)))+
  scale_x_continuous(breaks = seq(1,3,1))+
  #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
  xlab("Visit")+
  ylab('ng/ml')+
  ggtitle('Kynurenic.Acid')+
  theme_classic()+
  labs(color = 'Subject')
  #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
  #print(p)

edataName_visits = paste('Neopterin',c('V1','V2','V3'),sep='.')
clusters <- kmeans(trp_long[,edataName_visits],k)
cluster_trp$cluster = clusters$cluster
dat_cluster = merge(mx3,cluster_trp[,c(1,35)],by = 'Subject')
dat_cluster$nVisit = as.numeric(substring(dat_cluster$Visit,2,2))
#spaghetti plot of bile acids by clusters
p6 = ggplot(dat_cluster)+
  facet_grid(.~cluster)+
  geom_line(aes(nVisit,Neopterin,color = as.factor(Subject)))+
  scale_x_continuous(breaks = seq(1,3,1))+
  #geom_text(data = dat_ba1[dat_ba1$nVisit==2,],aes(x = 2,y = get(edataName),label = Subject))+
  xlab("Visit")+
  ylab('ng/ml')+
  ggtitle('Neopterin')+
  theme_classic()+
  labs(color = 'Subject')
  #ggsave(sprintf('%s/Report/plots/trajectory/%s_spaghetti.png',dir2,edataName),width = 6,height = 4)
  #print(p)

library(gridExtra)
grid.arrange(p1,p2,p3,p4,p5,p6,ncol = 2)


```