---
title: "Mother-baby microbiome longitudinal study"
author:
- Wanting Jin
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



# Outline of the analysis
* Using quantile regression to impute missing value due to limit of detection
* Delete HCA,MDCA, wMCA in bile acid because of high missing rate
* Transform data by log to normalize the data
* Boxplot and histogram to analyze the distribution and outliers
* Summary statistics to summarize the data
* Correlation heatmap between variables
* Association analysis by linear mixed model between bile acids and tryptophan/diversity/mood/psychiatric history




# Bile Acids analysis

## Boxplot and histogram for Bile Acids after log transformation

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}
### summary stat analysis for the Bile Acids
dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'
setwd(dir1)

library(knitr)
library(ggplot2)
library(reshape2)
library(lme4)
source('data_analysis_MBML.R')
library(tidyr)
library(stringr)
library(dplyr)
library(lmerTest)
library(magrittr)
library(qwraps2)

#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results/',dir2)


filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)

edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')


par(mfrow=c(3,3))
for(i in 1:length(edataNames)){
  edataName = edataNames[i]
  boxplot(dat_ba[,edataName], notch = FALSE, main=edataName, ylab=paste('ng/mL'))
  #boxplot.with.outlier.label(dat_trp[,edataName],dat_trp$sampID)
}

```


##box cox transformation for bile acids

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}
### summary stat analysis for the Bile Acids
dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'
setwd(dir1)
library(MASS)


filein_ba = sprintf('%s/imputed_bilacid.csv',dir_mb)
dat_ba = read.csv(filein_ba)
lambda = c()
for(i in 1:17){
  bc = boxcox(dat_ba[,i+2]~1,optimise = T)
  lambda[i]=bc$x[which.max(bc$y)]

}


edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')


```



## Boxplot of Bile Acids after log transformation by visits

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}

par(mfrow=c(3,3))
for(i in 1:length(edataNames)){
  edataName = edataNames[i]
  
  fm1 = sprintf('%s ~ Visit',edataName)
  fm1 = as.formula(fm1)
  boxplot(fm1, data= dat_ba, notch = FALSE,  main=edataName, ylab=paste('ng/mL'))
}

```

## Histogram of Bile Acids after log transformation

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*3, fig.height=3*3}

par(mfrow=c(3,3))
for(i in 1:length(edataNames)){
  edataName = edataNames[i]
  hist(dat_ba[,edataName], main=edataName, xlab=paste('ng/mL'),breaks=20)
  #boxplot.with.outlier.label(dat_trp[,edataName],dat_trp$sampID)
}


```

\newpage

## Summary statistics for Bile Acids

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE}


options(qwraps2_markup = "markdown")

idx_stat = c(1,4,3)
our_summary2 <-
  with(dat_ba,
       list('TaMCA' = tab_summary(TaMCA)[idx_stat],
            'THCA' = tab_summary(THCA)[idx_stat],
            'GHCA' = tab_summary(GHCA)[idx_stat],
            'TUDCA' = tab_summary(TUDCA)[idx_stat],
            'GUDCA' = tab_summary(GUDCA)[idx_stat], 
            'TCA' = tab_summary(TCA)[idx_stat],
            'GCA' = tab_summary(GCA)[idx_stat],
            'CA' = tab_summary(CA)[idx_stat],
            'UDCA' = tab_summary(UDCA)[idx_stat],
            'TCDCA' = tab_summary(TCDCA)[idx_stat],
            'GCDCA' = tab_summary(GCDCA)[idx_stat],
            'TDCA' = tab_summary(TDCA)[idx_stat],
            'GDCA' = tab_summary(GDCA)[idx_stat],
            'CDCA' = tab_summary(CDCA)[idx_stat],
            'DCA' = tab_summary(DCA)[idx_stat],
            'GLCA' = tab_summary(GLCA)[idx_stat],
            'bMCA' = tab_summary(bMCA)[idx_stat]
           #                    "mean (95% CI)" = ~ frmtci(qwraps2::mean_ci(Kynurenine, na_rm=TRUE)))
           # "Forward Gears" = tab_summary(as.character(gear))
       ))
            
            
whole <- summary_table(dat_ba, our_summary2)
#print(kable(whole))

grouped <- summary_table(dplyr::group_by(dat_ba, Visit), our_summary2)
#grouped

both = cbind(whole, grouped)
colnames(both)[1] = 'All'
print(both,rtitle='Bile Acids',font_size = 2,caption='Summary statistics for Bile Acids analysis')



```

\newpage


## Outlier detection for Bile Acids

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE}
### print the outlier subject, visit and values

### remove extreme subject values for other pain ratings under flavor = 0 and nico = 0
#painNames = c('THRES_COLD','TOLR_COLD','INTS_COLD','UNPLS_COLD','THRES_ISCH','TOLR_ISCH','INTS4_ISCH','UNPLS4_ISCH','MXNUM_ISCH')
### the extreme outliers are defined as:
# Extreme Outliers fall greater than 3.0*IQR outside the IQR and are plotted with a closed circle ['Tag'='OutliersX'].
# For example, if the lower and upper ranges on the IQR are 4 and 6, the IQR=2.0 and values between 6+(1.5*2) and 6+(3*2) ( = 9 and 12) 
# or between 4-(1.5*2) and 4-(3*2) ( = 1 and -2) are considered standard outliers. Samples above 12 or below -2 are considered extreme outliers.
 
mx1 = dat_ba

IQR_upper = rep(NA,length(edataNames))
for(j in 1:length(edataNames)){
	## I think we only need to worry about the upper extreme outlier in this study
	y1 = mx1[,edataNames[j]]
	IQR_upper[j] = quantile(y1, 3/4,na.rm=TRUE) + 3*IQR(y1, na.rm=TRUE)
	#boxplot_list[[i]] = boxplot(dat1[dat1$flavor == 0 & dat1$nico == 0,painNames[j]])
}


mx_ot = c()
#pain_rm = c()
for(i in 1:dim(mx1)[1]){
	for(j in 1:length(edataNames)){
		#if(dat1$flavor[i] == 0 & dat1$nico[i] == 0){
	  val = mx1[i,edataNames[j]]
	  #print(paste(val,IQR_upper[j]))
			if(!is.na(val) & val > IQR_upper[j]){
				#dat1[i,painNames[j]] = 'NA'
			  rowLine = c(mx1$Subject[i], mx1$Visit[i],edataNames[j],IQR_upper[j], val)
			  
				mx_ot = rbind(mx_ot, rowLine)
				#sub_rm = c(sub_rm,mx1$SUBJECT[i])
				#pain_rm = c(pain_rm,edataNames[j])
				

		  }
		#}
	}
}

colnames(mx_ot) = c('Subject','Visit','Measures','Upper range (Q3+3*IQR)','Value')

kable(mx_ot,row.names=FALSE)




```

\newpage


## Spaghetti plot for Bile Acids after log transformation

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=4*3, fig.height=4*3}
#interaction.plot(dat_trp$Visit, dat_trp$Subject, dat_trp$Tryptophan)

par(mfrow=c(3,3))
for(i in 1:length(edataNames)){
  edataName = edataNames[i]
  
  interaction.plot(dat_ba$Visit, dat_ba$Subject, dat_ba[,edataName], xlab="Visits", ylab='ng/mL', main=edataName, col=c(1:30), legend=F)
}

```



```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE, fig.height=2*6, fig.width=2*5}
#interaction.plot(dat_trp$Visit, dat_trp$Subject, dat_trp$Tryptophan)

par(mfrow=c(6,5))

for(i in 1:length(edataNames)){
  edataName = edataNames[i]
  ymax = max(dat_ba[,edataName],na.rm= TRUE)
  for(id in sort(unique(dat_ba$Subject))){
      dat1 = dat_ba[dat_ba$Subject == id,]
      dat1$Visit = as.integer(gsub('V','',dat1$Visit))
      plot(dat1$Visit, dat1[,edataName],xlab="Visits", ylab='ng/mL', main=paste(edataName,':','ID',id),type='l',ylim=c(0,ymax) )
  }

}
```


\newpage

## Correlation of Bile Acids after log transformation
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE, fig.height=6, fig.width=2*5}

library(ggplot2)
library(reshape2)

cormat = round(cor(dat_ba[,edataNames], use="pairwise.complete.obs"), 2)


melted_cormat <- melt(cormat)
#head(melted_cormat)



# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
upper_tri <- get_upper_tri(cormat)


reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

#Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white",
   midpoint = 0, limit = c(-1,1), space = "Lab",
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1,
    size = 12, hjust = 1))+
 coord_fixed()
# Print the heatmap
#print(ggheatmap)


ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))



```


\newpage

## Association analysis for Bile Acids

### Bile Acids of three visits as function of beta and alpha diversity of microbiome

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}

dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'


setwd(dir1)


library(knitr)
library("lme4")
library(lmerTest)
#library(survival)
#library(coxme)
source('data_analysis_MBML.R')





#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results',dir2)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)

filein_trp = sprintf('%s/mbml_trp.csv',dir_mb)
dat_trp = read.csv(filein_trp)
mx2 = merge(dat_trp[,c(1:5,26:47)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)


edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')




dataset = 'BA'






cvrtNames = 'Visit'
randomName = 'Subject'


  
  cdataNames = c(paste("wunifrac.PC.",1:4,sep='')
                  ,"chao1","observed_species","PD_whole_tree","shannon")
  

  dataName = sprintf('%s_vs_div',dataset)

  
  # Linear-mixed effect longitudinal model of random intercept and slope
  pairwise_association_test_lmm(mx2, cdataNames, edataNames, randomName, cvrtNames,dataName = dataName, dir_out = dir_out)
  


```




```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  
  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```




### Change of Tryptophan of three visits as function of V1/V2 of beta and alpha diversity of microbiome

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}

setwd(dir1)
library(lmerTest)

source('data_analysis_MBML.R')


#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results',dir2)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)

filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
dat_trp = read.csv(filein_trp)
mx2 = merge(dat_trp[,c(1:5,26:47)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)


edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')




dataset = 'BA'






cvrtNames = 'Visit'
randomName = 'Subject'


  
  cdataNames = c(paste("wunifrac.PC.",1:4,sep='')
                  ,"alphadiv.PC.1","alphadiv.PC.2")
  
  
  
  
  library(reshape2)
  #test1 = dcast(melt(mx2,id.vars=c('Subject','Visit')), Subject~variable+Visit)
  
  ## reshape the responses data to wide format
  mx2_long = reshape(mx2, idvar='Subject', timevar = 'Visit', direction='wide')
  
  edataNames_diff21 = c()
  edataNames_diff32 = c()
  
  # calculate the diff of Trp pathway variables
  for(edataName in edataNames){
    edataName_visits = paste(edataName,c('V1','V2','V3'),sep='.')
    edataName_diff21 = paste(edataName,'V21',sep='.')
    edataName_diff32 = paste(edataName,'V32',sep='.')
    mx2_long[[edataName_diff21]] = mx2_long[, edataName_visits[2]] - mx2_long[, edataName_visits[1]]
    mx2_long[[edataName_diff32]] = mx2_long[, edataName_visits[3]] - mx2_long[, edataName_visits[2]]
    
    edataNames_diff21 = c(edataNames_diff21, edataName_diff21)
    edataNames_diff32 = c(edataNames_diff32, edataName_diff32)

  }
  
  # association between div v1 and change of Trp
  dataName = sprintf('%s_diff_vs_div_v1',dataset)
  cdataNames_v1 = paste(cdataNames, 'V1',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)
  
  
    # association between div v2 and change of Trp
  dataName = sprintf('%s_diff_vs_div_v2',dataset)
  cdataNames_v1 = paste(cdataNames, 'V2',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)

```

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  
  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```



\newpage

### Mood assessment as function of Bile Acids

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*6}



setwd(dir1)


library(knitr)
library("lme4")
library(lmerTest)
#library(survival)
#library(coxme)



source('data_analysis_MBML.R')


#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results',dir2)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)

filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
dat_trp = read.csv(filein_trp)
mx2 = merge(dat_trp[,c(1:5,26:47)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)

  

cdataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')
  
edataNames = c('GAD','EPSD')

dataset = 'Mood'






cvrtNames = 'Visit'
randomName = 'Subject'


  

  dataName = sprintf('%s_vs_ba',dataset)

  pairwise_association_test_lmm(mx2, cdataNames, edataNames, randomName, cvrtNames,dataName = dataName, dir_out = dir_out)
  
```

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}

  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```






### Change of mood of three visits as function of V1/V2 of Bile Acids

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}



setwd(dir1)


library(knitr)
library("lme4")
#library(survival)
#library(coxme)



source('data_analysis_MBML.R')


#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results',dir2)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)

filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
dat_trp = read.csv(filein_trp)
mx2 = merge(dat_trp[,c(1:5,26:47)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)


edataNames = c('GAD','EPSD')
cdataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')
  




dataset = 'mood'






cvrtNames = 'Visit'
randomName = 'Subject'


  

  
  library(reshape2)
  #test1 = dcast(melt(mx2,id.vars=c('Subject','Visit')), Subject~variable+Visit)
  
  ## reshape the responses data to wide format
  mx2_long = reshape(mx2, idvar='Subject', timevar = 'Visit', direction='wide')
  
  edataNames_diff21 = c()
  edataNames_diff32 = c()
  
  # calculate the diff of Trp pathway variables
  for(edataName in edataNames){
    edataName_visits = paste(edataName,c('V1','V2','V3'),sep='.')
    edataName_diff21 = paste(edataName,'V21',sep='.')
    edataName_diff32 = paste(edataName,'V32',sep='.')
    mx2_long[[edataName_diff21]] = mx2_long[, edataName_visits[2]] - mx2_long[, edataName_visits[1]]
    mx2_long[[edataName_diff32]] = mx2_long[, edataName_visits[3]] - mx2_long[, edataName_visits[2]]
    
    edataNames_diff21 = c(edataNames_diff21, edataName_diff21)
    edataNames_diff32 = c(edataNames_diff32, edataName_diff32)

  }
  
  # association between div v1 and change of Trp
  dataName = sprintf('%s_diff_vs_ba_v1',dataset)
  cdataNames_v1 = paste(cdataNames, 'V1',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)
  
  
    # association between div v2 and change of Trp
  dataName = sprintf('%s_diff_vs_ba_v2',dataset)
  cdataNames_v1 = paste(cdataNames, 'V2',sep='.')

  pairwise_association_test(mx2_long, cdataNames_v1, c(edataNames_diff21, edataNames_diff32),dataName = dataName, dir_out = dir_out)

```

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=3*4, fig.height=3*5}
  
  timeName = 'Visit'
  
  par(mfrow=c(5,4))
  plot_longitudinal(mx2,cdataNames,edataNames, timeName,dataName = dataName, dir_out = dir_out)


```




### Tryptophan/Bile Acids of three visits as function of Psychiatric history

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE,warnings=FALSE,fig.width=2.5*4, fig.height=2.5*6}

dir1 = 'E:/Github/mbml/Scripts'
dir2 = 'E:/Wanting/mbml'

setwd(dir1)

library(knitr)
library("lme4")
library(lmerTest)
#library(survival)
#library(coxme)

source('data_analysis_MBML.R')


#dir_beha = 'behaviordata'
dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_out = sprintf('%s/results',dir2)

filein_ba = sprintf('%s/imputed_bilacid_log.csv',dir_mb)
dat_ba = read.csv(filein_ba)
filein_trp = sprintf('%s/mbml_trp1.csv',dir_mb)
dat_trp = read.csv(filein_trp)
filein_div = sprintf('%s/qiime2_diversity.csv',dir_mb)
dat_div = read.csv(filein_div)
mx2 = merge(dat_trp[,c(1:5,44:51)],dat_ba,by=c('Subject','Visit'),sort = FALSE,all=TRUE)
mx2 = merge(mx2,dat_div,by='barcode')


filein_MPPD = sprintf('%s/MPPD_Data_Master_09-13-2020.csv',dir_mb)
MPPD = read.csv(filein_MPPD)
history = MPPD[,c(1,21:22,24,28,31:34)]

mx3 = merge(mx2,history,by.x = 'subjectID',by.y = 'SUBJID')
#cdataNames = c(paste("wunifrac.PC.",1:4,sep='')
#                ,"pielou_e","observed_otus","faith_pd","shannon")
cdataNames = colnames(history[,3:9])

edataNames = c('Tryptophan','Kynurenine','Kynurenic.Acid','Neopterin','Serotonin', 'KYN_TRP_Ratio', 'KA_KYN_Ratio')
dataset = 'Trp'
cvrtNames = 'Visit'
randomName = 'Subject'
dataName = sprintf('%s_vs_his',dataset)
pairwise_association_test_lmm(mx3, cdataNames, edataNames, randomName, cvrtNames,dataName = dataName, dir_out = dir_out)


edataNames = c('TaMCA','THCA','GHCA','TUDCA'
               ,'GUDCA','TCA','GCA'
               ,'CA','UDCA','TCDCA','GCDCA','TDCA'
               ,'GDCA','CDCA','DCA','GLCA','bMCA')
dataset = 'BA'
dataName = sprintf('%s_vs_his',dataset)
pairwise_association_test_lmm(mx3, cdataNames, edataNames, randomName, cvrtNames,dataName = dataName, dir_out = dir_out)


edataNames = c(paste("wunifrac.PC.",1:4,sep='')
               ,"pielou_e","observed_otus","faith_pd","shannon")
dataset = 'div'
dataName = sprintf('%s_vs_his',dataset)
pairwise_association_test_lmm(mx3, cdataNames, edataNames, randomName, cvrtNames,dataName = dataName, dir_out = dir_out)

```

